\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-14 papers/itsc_2012/final.tex
section 3 and 4 revised based on comments

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \commentpk{cite a link to Scala here}. Scala provides the advantage of
functional programming constructs while still permitting imperative style.
Parts of AORTA, including extensions and clients, can easily be written in Java
or Scala. The software is
open-source\footnote{http://code.google.com/p/road-rage} and easily extensible
\commentpk{maybe "Extensions and clients built on top of AORTA can be written
in either Java or Scala, as the software is
open-source\footnote{\commentpk{needs more description}
http://code.google.com/p/road-rage} and easily extensible}. \commentpk {For
instance, the} The stop sign policy used in Section \ref{sec:results}'s
experiments \commentpk{Sec VI's experiments is awk, maybe "used in experiments
described in Sec VI"} is implemented in just $70$ \commentpk{remove \$\$ in
latex code, here and elsewhere} lines of code.

@@

One advantage of AORTA is that it simulates traffic on existing infrastructure
rather than contrived, randomly \commentpk{"contrived, randomly" is a bit
weird, maybe "contrived or randomly"?} generated maps. Utilizing OpenStreetMaps
\commentpk{remove trailing s in OpenStreetMaps}, a community project offering
user-editable, free XML-based maps, enables this. To run a simulation on a map
from OSM, there is a one-time conversion process to transform the map into
AORTA's format, which is also encoded with XML. \commentpk{why convert it to
another XML format? you might want to put an additional line outlining the
purpose of this step}

@@

The map also has a geometric interpretation, used for visualization and
\commentpk{measuring?} physical distances. A road contains a sequence of points
giving the line segments of the center yellow line \commentpk{if you want to
use "yellow", you need to reference a figure. I would recommend putting it this
way "center line of the road (yellow lines in Fig. XX)"}. These points could
also be interpreted as control points for a curved road, but AORTA uses
straight lines as an approximation. Lanes are parallel to the center line, some
fixed width apart from one another \commentpk{no need to mention that the roads
could be interpreted as a spline, instead say "AORTA interprets roads as
straight lines between these points, and projects parallel lanes out using a
fixed width"}. Turns are modeled with a single straight line connecting the end
of one lane to the start of another (so right turns often have $0$ length).

@@

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder \commentpk{you have not mentioned a builder before, maybe mention it in the last line of the intro to this section} marks GPS locations common to multiple roads, since
        they implicitly indicate intersections \commentpk{do you guys use GPS locations or OSM unique identifiers, what about overpasses etc.?}.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each direction
        (unless the road is marked one-way). The builder guesses the number of lanes based on
        OSM's ``road type'' \commentpk{If you have space, elaborate that you use heuristics on the road type attribute, and the 4 different road types are residential etc.} tag or an explicit number of lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        junction \commentpk{intersection, keep the terminology constant}. The builder uses an angle heuristic to choose between left, right, straight,
        and U-turns \commentpk{preceding sentence unclear to me}. When several lanes all cross into fewer lanes,
        the builder forces merging of the rightmost lanes.  Although inaccurate,
        these patterns are the best compromise. At the time of writing, even sources
        like Google Maps \cite{googleMaps} lack this information \commentpk{the last sentence is tricky. unless you have a source that states this, I would avoid this claim. Linking to google maps directly is not correct}.
  \item Because turns are heuristically placed, the graph is often
        disconnected. In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the digraph.
        Disconnected lanes are spliced out of larger roads, leaving unappealing
        geometric gaps, but correctly ensuring connectivity for corner-case free
        pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Simulation is discrete-time, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity, and
space-continuous, meaning roads are not tiled \commentpk{recommend breaking up
preceding sentence into 2. I lost track haflway}. The step duration $dt$ is
fixed, enabling agents to reason about collision avoidance and satisfying other
constraints \commentpk{preceding sentence is unclear}. Setting this value to
the average ``reaction time'' of a human driver may be appropriate for some
\commentpk{"human driver" instead of some?} experiments. Each step, the
simulation does the following:

@@

The speed of simulation depends on the step duration $dt$, the map size, and the
number of agents. A measure independent of these factors is the number of agent
steps per second. On a $2.4$ GHz machine, this can be observed to average at
about $150,000$, a number comparable to existing simulators \cite{SUMOthesis}.
When $dt = 0.1$ seconds, this means $15,000$ agents can be comfortably simulated
in real-time.  Further optimizations are planned.

@@

\begin{enumerate}
  \item the map (a $6$ by $7$ mile slice of downtown Austin, Texas
        \footnote{bounded by the Mopac and Springdale Road horizontally,
        and Oltorf Street and Koenig Lane vertically})
  \item $dt = 0.1$ seconds
  \item the agents spawned and their routes -- a continuous generator creates
        one new agent every $1$ simulation second, with a uniformly random
        starting position and goal
  \item the duration of the experiment: one hour
\end{enumerate}

@@

Each policy has parameters, so the configuration used is relevant. Stop signs
accept agents FCFS, refusing agents until they have idled at $0$ speed for $1$
second. This delay approximates human reaction time. The traffic signal policy
requires timings and groups of simultaneous turns for each intersection. The
configuration tested here uses a breadth-first search from one major road's
lanes to try to synchronize green signals for the agents near that lane to not
stop for several intersections. This heuristic needs improvement in how it
handles the remaining groups of turns that are scheduled after the first
sequence. Finally, the autonomous reservation policy allows new agents to join
existing groups of agents if their turn is compatible with the others. To
prevent one heavy direction of traffic from hogging the intersection
indefinitely, a timeout preemptively cycles through reservations. Ideally this
should be based on the relative numbers of agents depending on their
group of reservations.

@@

If an intersection immediately permits an agent to proceed and grants other
agents in the same lane similar access (because this new agent must still follow
these others), then the new agent may simply proceed at full speed limit without
pausing. This gives the least possible delay spent along some lane, so any extra
time is defined as lag and is used to measure intersection performance. Figure
\ref{fig:avg_lag} graphs the average of this delay per agent. On average,
reservation policies cause less than $1$ second of delay at each intersection,
stop signs cause about $7$ seconds lag, and traffic signals cause $16$ seconds.

@@

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. Figure
\ref{agent_cnt} reveals that autonomous reservations let agents reach their
destination twice as quickly as the competition, with a steady-state of about
$400$ agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.

Each policy has its trade-offs in terms of performance and physical hardware
cost, suggesting a mixed approach, perhaps with intersection policies adapting
as traffic conditions shift. Researchers could conduct this experiment and many
others without needing to understand most of AORTA's code-base. The largest
intersection policy, traffic signals with the flooding heuristic for timing
assignment, is only $400$ lines of Scala.

>>>>

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine \cite{scala}. Scala provides the advantage of functional programming
constructs while still permitting imperative style.  Extensions and clients
built on top of AORTA can be written in either Java or Scala. The software is
open-source \footnote{source code at http://code.google.com/p/road-rage} and
easily extensible. For instance, the stop sign sign policy used in experiments
described in section \ref{sec:result} is implemented in just 70 lines of code.

@@

One advantage of AORTA is that it simulates traffic on existing infrastructure
rather than randomly generated maps. Utilizing OpenStreetMap \cite{osm}, a
community project offering user-editable, free XML-based maps, enables this. To
run a simulation on a map from OSM, there is a one-time conversion process to
transform the map into AORTA's format, which encodes information about lanes and
turns not contained in the OSM source.

@@

The map also has a geometric interpretation, used for visualization and
measuring physical distances. A road contains a sequence of points, defining the
center line that divides the road into two directions of travel. AORTA
interprets these points as straight lines and projects parallel lanes out using
a fixed width. Turns are modeled with a single straight line connecting the end
of one lane to the start of another (so right turns often have 0 length).

@@

The builder works by incrementally transforming OSM graphs into AORTA graphs,
with each pass feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder marks GPS nodes common to multiple roads, since
        they implicitly indicate intersections. Overpasses do not share nodes
        with the roads they cross \cite{osmOverpass}.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each
        direction (unless the road is marked one-way). The builder guesses the
        number of lanes based on OSM's ``road type'' tag or an explicit number of
        lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        intersection. The angle between the two lanes determines whether the turn is
        a left, right, straight, or U-turn. When several lanes all cross into fewer
        lanes, the builder forces merging of the rightmost lanes. Although
        inaccurate, these patterns are the best compromise with the available data.
  \item Because turns are heuristically placed, the graph is often disconnected.
        In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the
        digraph.  Disconnected lanes are spliced out of larger roads, leaving
        unappealing geometric gaps, but correctly ensuring connectivity for
        corner-case free pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Simulation is discrete-time, meaning an agent accelerates at a fixed rate for
the duration of a ``step'' to achieve a new position and velocity, and
space-continuous, meaning roads are not tiled \commentpk{recommend breaking up
preceding sentence into 2. I lost track haflway}. The step duration dt is
fixed, enabling agents to reason about collision avoidance and satisfying other
constraints \commentpk{preceding sentence is unclear}. Setting this value to
the average ``reaction time'' of a human driver may be appropriate for some
\commentpk{"human driver" instead of some?} experiments. Each step, the
simulation does the following:

@@

The speed of simulation depends on the step duration dt, the map size, and the
number of agents. A measure independent of these factors is the number of agent
steps per second. On a 2.4 GHz machine, this can be observed to average at
about 150,000, a number comparable to existing simulators \cite{SUMOthesis}.
When dt = 0.1 seconds, this means 15,000 agents can be comfortably simulated
in real-time.  Further optimizations are planned.

@@

\begin{enumerate}
  \item the map (a 6 by 7 mile slice of downtown Austin, Texas
        \footnote{bounded by the Mopac and Springdale Road horizontally,
        and Oltorf Street and Koenig Lane vertically})
  \item dt = 0.1 seconds
  \item the agents spawned and their routes -- a continuous generator creates
        one new agent every 1 simulation second, with a uniformly random
        starting position and goal
  \item the duration of the experiment: one hour
\end{enumerate}

@@

Each policy has parameters, so the configuration used is relevant. Stop signs
accept agents FCFS, refusing agents until they have idled at 0 speed for 1
second. This delay approximates human reaction time. The traffic signal policy
requires timings and groups of simultaneous turns for each intersection. The
configuration tested here uses a breadth-first search from one major road's
lanes to try to synchronize green signals for the agents near that lane to not
stop for several intersections. This heuristic needs improvement in how it
handles the remaining groups of turns that are scheduled after the first
sequence. Finally, the autonomous reservation policy allows new agents to join
existing groups of agents if their turn is compatible with the others. To
prevent one heavy direction of traffic from hogging the intersection
indefinitely, a timeout preemptively cycles through reservations. Ideally this
should be based on the relative numbers of agents depending on their
group of reservations.

@@

If an intersection immediately permits an agent to proceed and grants other
agents in the same lane similar access (because this new agent must still follow
these others), then the new agent may simply proceed at full speed limit without
pausing. This gives the least possible delay spent along some lane, so any extra
time is defined as lag and is used to measure intersection performance. Figure
\ref{fig:avg_lag} graphs the average of this delay per agent. On average,
reservation policies cause less than 1 second of delay at each intersection,
stop signs cause about 7 seconds lag, and traffic signals cause 16 seconds.

@@

The number of agents changes because the continuous generator introduces a new
agent every second. This count reaches a steady state when the generator
introduces an agent at the same rate as older agents finish their route. Figure
\ref{agent_cnt} reveals that autonomous reservations let agents reach their
destination twice as quickly as the competition, with a steady-state of about
400 agents. The count for traffic signals continues to increase because
gridlock occurred in one segment of the map, causing agents to not finish their
route. Unexpected starvation was also observed: although the signals cycle
through turns regularly, agents cannot proceed when the previous cycle caused a
destination lane to fill and become blocked.

Each policy has its trade-offs in terms of performance and physical hardware
cost, suggesting a mixed approach, perhaps with intersection policies adapting
as traffic conditions shift. Researchers could conduct this experiment and many
others without needing to understand most of AORTA's code-base. The largest
intersection policy, traffic signals with the flooding heuristic for timing
assignment, is only 400 lines of Scala.
